---
title: "SEM Example Runs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SEM Example Runs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This tutorial shows how to set up SEM for a basic run and how to change the parameter values. 

## Set Up 

```{r setup}
library(PestED)
library(ggplot2)
```

Import and format the meteorological inputs. 

```{r}
# Define the path to the netcdf file containing the meteorology inputs.
nc_path <- system.file('extdata/AMF_USMe2_2005_L2_GF_V006.nc', package = 'PestED') 

# Format the inputs into the SEM required format. 
met_inputs <- format_inputs(nc_path)
```

## Basic Run

Now that the inputs are set up let's run SEM. 
```{r}
baseline <- iterate.SEM(pest = c(0, 0, 0, 0, 0), #  Disturb none of the carbon pools. 
            inputs = met_inputs, # Read in the met inputs. 
            params = PestED::default_parameters, # the list of default parameters is saved as package data. 
            years = .5) # the number of years to run SEM for. 
baseline$index <- 1:nrow(baseline) # add a time inndex to the results 
```

Let's take a quick  look at the results. 
```{r, fig.width=5, fig.height=5}
plot(baseline$GPP, ylab = 'GPP', main = 'Baseline SEM', type = 'l')
```

## Change Parameter Value

What happens when we divide Vcmax in half? 

```{r}
# Save a copy of the default parameter values. 
new_params <- PestED::default_parameters

# Alter the Vcmax value. 
new_params$Vcmax <- PestED::default_parameters$Vcmax / 2
```

Solve SEM with the new parameter value. 

```{r}
run2 <- iterate.SEM(pest = c(0, 0, 0, 0, 0), #  Disturb none of the carbon pools. 
            inputs = met_inputs, # Read in the met inputs. 
            params = new_params, # the list of default parameters is saved as package data. 
            years = .5) # the number of years to run SEM for. 
run2$index <- 1:nrow(run2) # add a time inndex to the results, to do is to add actual time inforamtion.
```

Compare the GPP of the two runs. 

```{r, fig.width=5, fig.height=5}
ggplot() + 
  geom_line(data = baseline, aes(index, GPP, color = 'baseline')) + 
    geom_line(data = run2, aes(index, GPP, color = 'lower Vcmax')) + 
  theme_bw() + 
  labs(title = 'GPP from two SEM runs', 
       x = 'Time Index', 
       y = 'GPP')
```
